Subject-level developmental relationships

### 2024-09-26 1_get_data.qmd
more paper stuff, CFA

```{r}
d_sub <- d_trial |>
  group_by(dataset_name, subject_id, administration_id, age) |>
  summarise(across(c(contains("_window"),"rt","log_rt"), ~mean(.x, na.rm=TRUE)))
```

>>>>

```{r}
d_sub <- d_trial |>
  group_by(dataset_name, subject_id, administration_id, age) |>
  summarise(n_trials = sum(!is.na(long_window_accuracy)), 
            n_trials_rt = sum(!is.na(rt)),
            across(c(contains("_window"),"rt","log_rt"), ~mean(.x, na.rm=TRUE)))
```

### 2024-09-26 2_developmental_analysis.qmd
more paper stuff, CFA

```{r}
d_var <- d_trial |>
  group_by(dataset_name, administration_id, age) |>
  summarise (`SD(log RT)` = sd(log_rt, na.rm=TRUE), 
             `SD(accuracy)` = sd(long_window_elogit, na.rm=TRUE)) |>
  pivot_longer(contains("sd"), 
               values_to = "sd", names_to = "measure")

v <- ggplot(d_var, aes(x = age, y = sd)) + 
  geom_point(alpha = .05) + 
  geom_smooth() +
  geom_smooth(method = "lm", col = "green") +
  geom_smooth(data = d_var |>
  filter(dataset_name %in% datasets_with_age_variation$dataset_name),
              aes(group = dataset_name, col = dataset_name), geom="line",
              alpha = .1, method = "lm", size = .5, lty = 2) + 
  scale_color_discrete(guide = FALSE) + 
  scale_x_log10() +
  facet_wrap(~measure, scales = "free_y")

@@

```{r}
d_sub |>
  group_by(dataset_name) |>
  summarise(n = n(), 
            `Mean Age` = mean(age), 
            `Min Age` = min(age), 
            `Max Age` = max(age), 
            `CDIs` = ifelse(!is.na(any(prod)), "x", ""), 
            `longitudinal` = ifelse(any(duplicated(subject_id)), "x", "")) |>
  xtable(caption = "Characteristics of included datasets from Peekbank.")
```

```{r}
a <- ggplot(d_sub_long_elogit |> filter(measure == "long_window_elogit"), 
       aes(x = age, y = accuracy)) + 
  geom_point(alpha = .05) + 
  geom_smooth() +
  geom_smooth(method = "lm", col = "green") +
  geom_smooth(data = d_sub_agevar_long_elogit |> filter(measure == "long_window_elogit"),
              aes(group = dataset_name, col = dataset_name), geom="line",
              alpha = .1, method = "lm", size = .5, lty = 2) + 
  geom_hline(yintercept = .5, lty = 2) + 
  scale_x_log10() +
  scale_color_discrete(guide = FALSE) + 
  ylab("Accuracy (elogit)") + 
  xlab("Age (months)")

>>>>

```{r}
d_var <- d_trial |>
  group_by(dataset_name, administration_id, age) |>
  summarise (`log(RT)` = sd(log_rt, na.rm=TRUE), 
             `Accuracy` = sd(long_window_accuracy, na.rm=TRUE)) |>
  pivot_longer(c("log(RT)","Accuracy"), 
               values_to = "sd", names_to = "measure")

v <- ggplot(d_var, aes(x = age, y = sd)) + 
  geom_point(alpha = .05) + 
  geom_smooth() +
  geom_smooth(method = "lm", col = "green") +
  geom_smooth(data = d_var |>
  filter(dataset_name %in% datasets_with_age_variation$dataset_name),
              aes(group = dataset_name, col = dataset_name), geom="line",
              alpha = .1, method = "lm", size = .5, lty = 2) + 
  scale_color_discrete(guide = FALSE) + 
  scale_x_log10() +
  facet_wrap(~measure, scales = "free_y") + 
  ylab("Standard Deviation") + 
  xlab("Age (months)")

@@

```{r}
d_sub |>
  group_by(dataset_name) |>
  summarise(`N subjects` = length(unique(subject_id)),
            `N admins` = n(), 
            `Mean Age` = mean(age), 
            `Min Age` = min(age), 
            `Max Age` = max(age), 
            `Avg Trials` = mean(n_trials),
            `Avg RT Trials` = mean(n_trials_rt),
            `CDIs` = ifelse(!is.na(any(prod)), "x", ""), 
            `longitudinal` = ifelse(any(duplicated(subject_id)), "x", "")) |>
  arrange(desc(`N subjects`)) |>
  xtable(caption = "Characteristics of included datasets from Peekbank. `Admins` denotes separate experimental sessions.")
```

```{r}
a <- ggplot(d_sub_long |> filter(measure == "long_window_accuracy"), 
       aes(x = age, y = accuracy)) + 
  geom_point(alpha = .05) + 
  geom_smooth() +
  geom_smooth(method = "lm", col = "green") +
  geom_smooth(data = d_sub_agevar_long |> filter(measure == "long_window_accuracy"),
              aes(group = dataset_name, col = dataset_name), geom="line",
              alpha = .1, method = "lm", size = .5, lty = 2) + 
  geom_hline(yintercept = .5, lty = 2) + 
  scale_x_log10() +
  scale_color_discrete(guide = FALSE) + 
  ylab("Accuracy") + 
  xlab("Age (months)")

### 2024-09-26 3_developmental_relationships.qmd
more paper stuff, CFA

```{r}
suppressPackageStartupMessages(library(here))
suppressPackageStartupMessages(library(psych))
suppressPackageStartupMessages(source(here("helper","common.R")))
d_sub <- readRDS(here("cached_intermediates","1_d_sub.Rds"))
```

@@

@@

d_sub_mat_grp <- d_sub_mat_dataset |>
  nest() |>
  mutate(complete_data = map(data, ~filter(.x[complete.cases(.x),])),
    mat = as.matrix(complete_data), 
    has_data = map_lgl(complete_data, ~nrow(.x)>0)) |>
  filter(has_data) |>
  mutate(factors = map(mat, ~fa(.x, 1, rotate = "varimax")), 
         f_summary = map(factors, summary), 
         loadings = map_df(factors, ~as.tibble(loadings(x))))

d_sub_mat_grp |>

@@

@@

d_sub_long <- d_sub |>
  filter(subject_id %in% longitudinal$subject_id, 
                     !is.na(prod)) |>
  group_by(subject_id) |>
  arrange(administration_id) |>
  mutate(admin_num = 1:n(), 
         delta_t = c(0,diff(age)))

ggplot(filter(d_sub_long, admin_num > 1), 
       aes(x = delta_t)) +
  geom_histogram()
```

```{r}
library(lavaan)

>>>>

```{r}
suppressPackageStartupMessages(library(here))
suppressPackageStartupMessages(library(psych))
suppressPackageStartupMessages(source(here("helper","common.R")))
d_sub <- readRDS(here("cached_intermediates","1_d_sub.Rds"))
library(lavaan)
library(tidySEM)

@@

## All data

```{r}
d_sub_mat <- d_sub |>
  ungroup() |>
  select(log_age, log_rt, long_window_accuracy)

fa.parallel(d_sub_mat)
```
```{r}
mod1 <- fa(d_sub_mat, 1, rotate = "varimax")

@@

d_sub_mat_grp <- d_sub_mat_dataset |>
  nest() |>
  mutate(complete_data = map(data, ~filter(.x[complete.cases(.x),])),
    mat = as.matrix(complete_data), 
    has_data = map_lgl(complete_data, ~nrow(.x)>0)) |>
  filter(has_data) |>
  mutate(fas = map(mat, ~fa(.x, 1, rotate = "varimax")), 
         f_summary = map(fas, summary), 
         loadings = map(fas, ~tibble(loadings = as.double(loadings(.x)), 
                                     name = rownames(loadings(.x)))))

ls <- d_sub_mat_grp |>
  select(dataset_name, loadings) |>
  unnest(cols = c(loadings))

ggplot(ls, aes(x = name, y = loadings, col = dataset_name)) + 
  geom_point(aes = .5) +
  stat_summary(fun.data = "mean_cl_boot", aes(group = 1)) +
  geom_hline(yintercept = 0, lty = 2) + 
  coord_flip()

@@

# CFA models

d_cfa <- d_sub |>
             rename(`accuracy` = long_window_accuracy) |>
             mutate(accuracy = scale(accuracy), 
                    log_age = scale(log_age), 
                    log_rt = scale(log_rt)) |>
  select(accuracy, log_age, log_rt)

d_cfa <- d_cfa[complete.cases(d_cfa), ]

fa_model <-  "F1  =~ log_age + log_rt + accuracy"
fit <- cfa(fa_model, d_cfa, std.lv=TRUE)

summary(fit)
a <- graph_sem(model = fit)+ theme(panel.background = element_rect(fill = "white"))
```

```{r}
d_cfa_prod <- d_sub |>
             rename(`accuracy` = long_window_accuracy) |>
             mutate(accuracy = scale(accuracy), 
                    log_age = scale(log_age), 
                    log_rt = scale(log_rt), 
                    prod = scale(prod)) |>
  select(accuracy, log_age, log_rt, prod)

d_cfa_prod <- d_cfa_prod[complete.cases(d_cfa_prod), ]

fa_model_prod <-  "F1  =~ log_age + log_rt + accuracy + prod"
fit_prod <- cfa(fa_model_prod, d_cfa_prod, std.lv=TRUE)

summary(fit_prod)
b <- graph_sem(model = fit_prod) + 
  theme(panel.background = element_rect(fill = "white"))
b
```
```{r}
d_cfa_comp <- d_sub |>
             rename(`accuracy` = long_window_accuracy) |>
             mutate(accuracy = scale(accuracy), 
                    log_age = scale(log_age), 
                    log_rt = scale(log_rt), 
                    comp = scale(comp)) |>
  select(accuracy, log_age, log_rt, comp)

d_cfa_comp <- d_cfa_comp[complete.cases(d_cfa_comp), ]

fa_model_comp <-  "F1  =~ log_age + log_rt + accuracy + comp"
fit_comp <- cfa(fa_model_comp, d_cfa_comp, std.lv=TRUE)

summary(fit_comp)
c <- graph_sem(model = fit_comp,
            layout = get_layout(x = fit_comp, layout_algorithm = "layout_as_star")) + 
    theme(panel.background = element_rect(fill = "white"))
c
```

```{r}
plot_grid(a,b,c, nrow = 1, labels = c("A","B","C"))
ggsave(here("figures","sem.png"), width = 25, height = 4)
```

@@

d_sub_long <- d_sub |>
  filter(subject_id %in% longitudinal$subject_id) |>
  group_by(subject_id) |>
  arrange(administration_id) |>
  mutate(admin_num = 1:n(), 
         delta_t = c(0,diff(age)))

ggplot(d_sub_long, 
       aes(x = delta_t, fill = as.factor(admin_num))) +
  geom_histogram() + 
  facet_wrap(~dataset_name)
```

Longitudinal data with CDI is just FMW2013 right now.

```{r}
fmw <- filter(d_sub_long, dataset_name == "fmw_2013") |>
  mutate(t = ifelse(age < 23, "t1","t2")) |>
  group_by(subject_id, t) |>
  summarise(age = mean(age), 
            log_rt = mean(log_rt), 
            prod = mean(prod), 
            long_window_accuracy = mean(long_window_accuracy),
            delta_t = mean(delta_t)) |>
  pivot_wider(names_from = "t", 
              values_from = c("age","log_rt","prod","long_window_accuracy", "delta_t"))

summary(lm(prod_t2 ~ long_window_accuracy_t1 * log_rt_t1 + age_t2, 
             data = fmw))
```
